DeepSphere: a graph-based spherical CNN

### 2019-09-24 main.tex
uneven sampling experiment

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
\end{itemize}
}
\newpage

@@

\begin{figure}[h]
	\centering
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_cosmo_cmb}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_brain_meg}
	} \hspace{0.1em}
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_climate_TMQ}
		\label{fig:examples:climate}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_daily_tmax}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_graph}
	}
	\caption{
		Examples of intrinsically spherical data:
		(a) the cosmic microwave background (CMB) temperature from \citet{planck2015overview},
		(b) brain activity recorded through magnetoencephalography (MEG),\protect\footnotemark
		(c) hourly precipitation from a climate simulation \citep{jiang2019sphericalcnn}, %\citep{mudigonda2017climateevents},
		(d) daily maximum temperature from the Global Historical Climatology Network (GHCN).\protect\footnotemark
		A rigid full-sphere pixelization is not ideal: the Milky Way's galactic plane masks observations, brain activity is only measured on the scalp, climate scientists desire a variable resolution, and the position of weather stations is arbitrary and changes over time.
		(e) Graphs can faithfully and efficiently represent sampled spherical data by placing vertices where data has been measured.
		%\todo{something from climate, weather?}
		%\todo{compare ideal full-spheres (like SHREC-17 projection from Esteves, 360 from Coors) to real-world measurements?}
		%\todo{horizontal colorbar for brain}
	}
	\label{fig:examples}
\end{figure}
\footnotetext[1]{\scriptsize\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}
\footnotetext[2]{\scriptsize\url{https://martinos.org/mne/stable/auto_tutorials/plot_visualize_evoked.html}}

@@

\subsection{GHCN} \label{sec:exp:ghcn}

* task is artificial, but it shows DeepSphere's flexibility\\
* structure (looking around) help for prediction\\
* equivariance to be verified in this case (next paper)\\

* non-regular sampling with different density over the globe (continent vs ocean)\\
* resulting graph is not rotation equivariant\\
* Two different tasks: a dense regression and a global regression\\

\paragraph*{dense regression - future temperature estimation}~\\
* Task is to find the temperature at day T, knowing the temperature of the 5 previous days.\\
* Goal is to change the order of the chebyshev filter to find the influence of the structure.\\
* order 0 is the same as time series, as the pixel have no information of its neighborhood.\\
* increasing order is good until threshold.\\
* The NN indeed predicts the correct temperature at day T, and not day T-1 (baseline) even though there is small difference.\\

\paragraph*{global regression - day in year estimation}~\\
* days represented by a sine to represent the periodic nature of the year.\\
* task too easy, temperature feature already periodic $\Rightarrow$ second experiment with only precipitation feature\\
* structure is still helping\\

\begin{table}
    \centering
    \begin{tabular}{c|ccc}
        order & future temp & day (temperature) & day (precipitations) \\ \hline
        0 & 0.896 & 0.881 & - 0.920\\
        4 & 0.919 & 0.969 & 0.597\\
    \end{tabular}
    \caption{R2 coefficient for different tasks on the GHCN dataset}
    \label{tab:GHCN_results}
\end{table}

@@

%We thank Taco Cohen for his inputs on the intriguing results of our comparisons with their gauge CNN \citep{}.

@@

\section{Proof of theorem}\label{sec: appendix: proof of theorem}
Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of a closed, compact and infinitely differentiable manifold $\mathcal M$, a smooth function $f:\mathcal M \rightarrow \mathbb R$, and defined the vector $\b f$ of samples of $f$ as follows: $\b f\in \mathbb R^n,\ \b f_i = f(x_i)$, it is known that  (\cite{belkin2005towards}) to prove the convergence of $(\b {L}_n^t \b f) _i$ to $\Delta_{\mathbb S^2}f(x_i)$ it is sufficient to prove the convergence of its continuous counterpart $L_n^tf(x)$ to $\Delta_{\mathbb S^2}f(x)$
where $L_n^t$ is defined as follows:
\begin{definition}{}\\
	\label{def:Heat Kernel Graph Laplacian operator}
	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=0}^{n-1} \exp \left\{ {-\frac{||x_i-y||^2}{4t}}\right\}\left(f(y)-f(x_i)\right)\right]$$
\end{definition}
Thus, to prove theorem \ref{theo:pointwise convergence for a regular sampling} it will be sufficient to prove that 
\begin{equation}\label{eq:continuous convergence}
 \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n^2} {L}_n^{t_n} f(y) =  \Delta_{\mathbb{S}^2}f(y)
\end{equation}
For the sake of the proof we will need also the following facts:
\begin{definition}{} (\cite{belkin2005towards})\\ \label{eq: my L^t} Let $\mu$ be the uniform probability measure on the manifold $\mathcal M$, and let $\text{vol}(\mathcal M)$ be the volume of $\mathcal M$. We define the functional approximation to the Laplace-Beltrami operator to be the operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\label{def:Functional approximation to the Laplace-Beltrami operator}
	$$ L^tf(y) = \int_{\mathcal M}\exp\left\{-\frac{||y-x||^2}{4t}\right\}\left(f(y)-f(x)\right)d\mu(x)$$
\end{definition}

@@

\begin{table}
    \centering
    \begin{tabular}{c|ccc}
        order & MSE & MAE & R2 \\ \hline
        0 & 10.88 & 2.42 & 0.896\\
        1 & 8.91 & 2.20 & 0.906\\
        4 & 8.20 & 2.11 & 0.919\\
        9 & 8.38 & 2.12 & 0.915\\
    \end{tabular}
    \caption{Future temperature estimation task. Evolution of performance in respect of the order of the graph filter}
    \label{tab:future_results}
    \hfill
    %~\\[1cm]
    % \begin{minipage}{0.48\linewidth}
    % \begin{tabular}{c|ccc}
    %     order & MSE & MAE & R2 \\ \hline
    %     1 & 134.01 & 10.51 & -0.319\\
    %     4 & 129.48 & 10.31 & -0.274\\
    %     9 & 141.68 & 10.82 & -0.389\\
    % \end{tabular}
    % \caption{Baseline: testing with the present day (T-1) instead of the future day T}
    % \label{tab:present_results}
    % \end{minipage}
\end{table}

\begin{table}
    \centering
    \begin{minipage}{0.48\linewidth}
    \begin{tabular}{c|ccc}
        order & MSE & MAE & R2 \\ \hline
        0 & 0.58 & 0.42 & -0.92\\
        4 & 0.50 & 0.18 & 0.597\\
    \end{tabular}
    \caption{Find day in year using precipitation}
    \label{tab:glob_prec}
    \end{minipage}
    \hfill
    %~\\[1cm]
    \begin{minipage}{0.48\linewidth}
    \begin{tabular}{c|ccc}
        order & MSE & MAE & R2 \\ \hline
        0 & 0.10 & 0.10 & 0.881\\
        4 & 0.05 & 0.05 & 0.969\\
    \end{tabular}
    \caption{Using temperature feature that is already periodical}
    \label{tab:glob_all}
    \end{minipage}
\end{table}

>>>>

\todo{
\begin{itemize}
	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN.
	\item adjacency matrix: $A$ instead of $W$?
	\item anonymize DeepSphere citation
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item citet for inline, citep for parenthesis
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method.
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot
	\item result tables: only what supports the main message in the text, everything we have in the appendix
	\item Sampling, discretization, or pixelization of the sphere? Be consistent.
\end{itemize}
}
\newpage

@@

\begin{figure}[h]
	\centering
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_cosmo_cmb}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_brain_meg}
	} \hspace{0.1em}
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_climate_TMQ}
		\label{fig:examples:climate}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_daily_tmax}
		\label{fig:examples:ghcn:tmax}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_graph}
		\label{fig:examples:ghcn:graph}
	}
	\caption{
		Examples of intrinsically spherical data:
		(a) the cosmic microwave background (CMB) temperature from \citet{planck2015overview},
		(b) brain activity recorded through magnetoencephalography (MEG),\protect\footnotemark
		(c) hourly precipitation from a climate simulation \citep{jiang2019sphericalcnn}, %\citep{mudigonda2017climateevents},
		(d) daily maximum temperature from the Global Historical Climatology Network (GHCN).\protect\footnotemark
		A rigid full-sphere pixelization is not ideal: the Milky Way's galactic plane masks observations, brain activity is only measured on the scalp, climate scientists desire a variable resolution, and the position of weather stations is arbitrary and changes over time.
		(e) Graphs can faithfully and efficiently represent sampled spherical data by placing vertices where data has been measured.
		%\todo{something from climate, weather?}
		%\todo{compare ideal full-spheres (like SHREC-17 projection from Esteves, 360 from Coors) to real-world measurements?}
		%\todo{horizontal colorbar for brain}
	}
	\label{fig:examples}
\end{figure}
\footnotetext[1]{\scriptsize\url{https://martinos.org/mne/stable/auto_tutorials/plot_visualize_evoked.html}}
\footnotetext[2]{\scriptsize\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}

@@

\subsection{Uneven sampling} \label{sec:exp:ghcn}

To demonstrate the flexibility of modeling the discretized sphere by a graph, we collected historical measurements from weather stations scattered across the Earth.\footnote{\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}
The spherical data is heavily non-uniformly sampled, with a much higher density of weather stations over North America than the Pacific (\figref{examples:ghcn:tmax}).
There are \todo{xx} stations in total.
For illustration, we devised two artificial tasks.
A dense regression that consists of predicting the temperature on a given day knowing the temperature on the 5 previous days.
A global regression that consists of predicting the day (represented as one period of a sine over the year) from temperature or precipitations.

\begin{table}
    \centering
    \begin{tabular}{
			c
			S[table-format=2.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=1.3]
			S[table-format=1.2]
			S[table-format=1.2]
			S[table-format=-1.3]
		}
		\toprule
		& \multicolumn{3}{c}{temp. (from past temp.)} & \multicolumn{3}{c}{day (from temperature)} & \multicolumn{3}{c}{day (from precipitations)} \\
		\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
		order $K$ & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} & {MSE} & {MAE} & {R2} \\
		\midrule
		$0$ & 10.88 & 2.42 & 0.896 & 0.10 & 0.10 & 0.881 & 0.58 & 0.42 & -0.920 \\
		%%$1$ &  8.91 & 2.20 & 0.906 & - & - & - & - & - & - \\
		$4$ &  8.20 & 2.11 & 0.919 & 0.05 & 0.06 & 0.969 & 0.50 & 0.18 &  0.597 \\
		%$9$ &  8.38 & 2.12 & 0.915 & - & - & - & - & - & - \\
		\bottomrule
    \end{tabular}
    \caption{
		Prediction results on data from weather stations.
		Structure always helps.
	}
    \label{tab:ghcn}
\end{table}

Results from \tabref{ghcn} show that treating the times series as a time-varying spherical signal (with filters of polynomial order $K>0$) yields higher performance than treating each time series independently ($K=0$).
Note that predicting the day from temperature is much easier as temperature has a yearly pattern.
Still, structure helps.

% * increasing order is good until threshold \\
% * equivariance to be verified in this case (next paper) -> in conclusion \\
% * The NN indeed predicts the correct temperature at day T, and not day T-1 (baseline) even though there is small difference. \\
% * \todo{performance of baseline} \\

@@

%We thank Pierre Vandergheynst for advices, and Andreas Loukas for having processed the GHCN data.
%We thank Taco Cohen for his inputs on the intriguing results of our comparison with \citet{cohen2019gauge}.
%We used the Python Graph Signal Processing package (PyGSP) \citep{pygsp} for computations and plots.
%The Python Graph Signal Processing package (PyGSP) \citep{pygsp} was used to build graphs, compute the Laplacian and Fourier basis, and perform graph convolutions.

@@

\section{Proof of theorem}\label{sec: appendix: proof of theorem}
Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of a closed, compact and infinitely differentiable manifold $\mathcal M$, a smooth function $f:\mathcal M \rightarrow \mathbb R$, and defined the vector $\b f$ of samples of $f$ as follows: $\b f\in \mathbb R^n,\ \b f_i = f(x_i)$, it is known that  (\cite{belkin2005towards}) to prove the convergence of $(\b {L}_n^t \b f) _i$ to $\Delta_{\mathbb S^2}f(x_i)$ it is sufficient to prove the convergence of its continuous counterpart $L_n^tf(x)$ to $\Delta_{\mathbb S^2}f(x)$
where $L_n^t$ is defined as follows:
\begin{definition}{}
	\label{def:Heat Kernel Graph Laplacian operator}
	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=0}^{n-1} \exp \left\{ {-\frac{||x_i-y||^2}{4t}}\right\}\left(f(y)-f(x_i)\right)\right]$$
\end{definition}
Thus, to prove theorem \ref{theo:pointwise convergence for a regular sampling} it will be sufficient to prove that 
\begin{equation}\label{eq:continuous convergence}
 \lim_{n\to\infty} \frac{|\mathbb{S}^2|}{4\pi t_n^2} {L}_n^{t_n} f(y) =  \Delta_{\mathbb{S}^2}f(y)
\end{equation}
For the sake of the proof we will need also the following facts:
\begin{definition}{} (\cite{belkin2005towards})\\ \label{eq: my L^t} Let $\mu$ be the uniform probability measure on the manifold $\mathcal M$, and let $\text{vol}(\mathcal M)$ be the volume of $\mathcal M$. We define the functional approximation to the Laplace-Beltrami operator to be the operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\label{def:Functional approximation to the Laplace-Beltrami operator}
	$$ L^tf(y) = \int_{\mathcal M}\exp\left\{-\frac{||y-x||^2}{4t}\right\}\left(f(y)-f(x)\right)d\mu(x)$$
\end{definition}

@@
